Title here

### 2016-12-16 portfolio/text/ms.Rmd

# Abstract

A fundamental challenge for individuals who rely on the extraction of natural
resources for their livelihood is income variability. Fishers, loggers, farmers,
and miners, among others, face high levels of income uncertainty driven by a mix
of environmental, biological, and economic variability. A key component
underlying the sustainability of these industries is identifying both how
regulatory actions and individual behaviour affect income variability and by
extension the economic stability and sustainable use of natural resources. For
commercial fisheries, recent work has identified that targeting a diversity of
species can reduce income variability. However, any diversity-variability
relationship is embedded in a regulatory framework that restricts individual
ability to change fishing behaviour. Here we identify three primary axes by
which fishers in Alaska, one of the world's most productive fishery regions, can
potentially diversify fishing activities and investigate how individual actions
along each diversity axis affect both revenue and revenue variability. We
construct a novel statistical model to simultaneously assess the value to
individual fishers of choosing to (1) own permits that allow targeting a diverse
fish species assemblage, (2) diversify by owning multiple permits, or (3) catch
a greater species diversity within the constraints of a given permit. We show
contrasting effects of diversity on revenue and revenue variability along the
three axes. First, we show that fishing permits with higher species diversity
reduces variability (`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction
per unit increase in mean effective species diversity). Next, we show that
fishing an additional permit is associated with higher revenue and lower
variability (e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`%
IQR] lower expected standard deviation for halibut plus another permit vs.
single halibut permits). Finally, we show that increasing species diversity
within the constraints of a permit set has a fishery-dependent effect on revenue
and is usually (`r g2_dens`% probability) associated with increased uncertainty
in revenue the following year. Thus, while there are clear benefits to some axes
of diversification, others are less effective. Importantly, the most effective
measure for increasing revenue and decreasing risk --- purchasing additional
permits to participate in additional fisheries --- was also the most costly. At
the same time, Alaskan commercial fisheries have been increasingly focussed on
single-permit (often salmon) fishing strategies. Our results suggest that the
benefit of species diversity for revenue variability is mostly an across-permit
effect, not a benefit that can be easily harnessed by an individual permit
holder. Effective methods of diversification --- purchasing additional permits
or switching permits --- are costly and often limited by regulations such as
limited entry programs and individual transferable quotas (ITQs). Take home ...

@@

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
obtainable by commercial fishers [@kasperski2013], possibly because both fish
and fishers are spatially mobile. However, fishers still have the ability to
lower risk. They can, for example, target as many species as permitted by
a license, own multiple licenses, or diversify spatially and fish in multiple
regions. The effectiveness of these diversification strategies for increasing
revenue and reducing revenue variability remains uncertain. For example, theory
and empirical studies suggest there can be substantial benefits to
specialization for both farmers and fishers [@abdulai2000; @arnason2005].
Furthermore, the majority of research examining the effects of diversification
on revenue do so at aggregate levels that combine individuals (e.g. households,
counties, or communities; [@kasperski2013; @sethi2014]). Because of this, key
uncertainties remain about the fundamental relationship between diversity,
fishing revenue, and fishing revenue variability at the level of the individual.

@@

Using a unique dataset collected over 30 years (1985--2014) on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals we can track
people that switch fishery permits over time or use multiple vessels. Similarly,
we can disentangle fishers that share the same vessel. We develop and apply
a hierarchical Bayesian model to understand the effect of fished species
diversity on individual revenue and revenue variability that builds on results
from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
[@sethi2014] variability.

Our model is a variance-function regression model [@davidian1987; @hall1989],
which simultaneously models changes in the mean and variance of individual
revenue as a function of covariates. Whereas traditional hierarchical linear
models assume homoscedasticity --- that the residual variance does not vary
systematically --- the variance-function regression model allows the residual
variance to be heteroscedastic and vary as a function of predictors. Our
specific model simultaneously estimates the effect of species diversity (the
inverse of Simpson's diversity index [@simpson1949; @jost2006]) on individual
revenue and revenue variability, and estimates effects both within and across
fishing strategies (sets of similar licenses or permits). At the same time, it
controls for fishing effort and latent time effects (e.g. changing environmental
conditions). To our knowledge, this is the first application of this type of
model in any natural resource field.

@@

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given permit set is more variable (Fig.
\ref{fig:within}). After accounting for fishing effort, specializing tends to be
associated with increased revenue (Fig. \ref{fig:within}; Fig. S\ref{fig:fe},
Fig. S\ref{fig:re}). On average, there is a ~0.5% increase in the year-to-year
revenue ratio per 1% decrease in species diversity, and across strategies, `r
g2_dens`% of the probability density supports a positive relationship between
specialization and revenue. The converse is also usually true --- generalizing
tends to be associated with lower expected revenue after accounting for fishing
effort (i.e. Fig. S\ref{fig:mirror}A and S\ref{fig:mirror}B are near mirror
images left to right). This benefit of specializing is most apparent after
accounting for fishing effort. Given that fishing for more species tends to be
associated with increased fishing effort (fishing more days), specializing tends
to be ...TODO (Fig. S\ref{fig:noeffort}).

After accounting for fishing effort, adjusting (increasing or decreasing) an
individual's fished species diversity within the constraints of a permit set
tends to be associated with greater expected variability or uncertainty in
year-to-year revenue (Fig. \ref{fig:within} strategies above 0; Figs.
S\ref{fig:fe}, \ref{fig:re}, \ref{fig:mirror}). This effect was strongest for
individuals with a halibut permit or halibut and sablefish permits. Individuals
with these strategies experienced ~2% increase in variability for a 1% change in
the year-to-year ratio of species diversity (Fig. \ref{fig:within}). Levels of
fished species diversity above one for these particular permits are primarily
due to bycatch.

Alaskan halibut and sablefish fisheries were converted to individual fishery
quota (IFQ) management in 1994. This catch share system... TODO. We refit our
models allowing all strategy-level effects to vary separately before and after the
introduction of IFQs for strategies involving halibut and sablefish permits. The
introduction of IFQs reduced variability by ~25% (XX--XX 95% CI) for individuals
with only halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for
most strategies, the association changing species diversity with increased
revenue variability became stronger after the introduction of IFQs (Fig.
S\ref{fig:ifq}B). This was especially true for individuals fishing only halibut
or sablefish and less so for individuals combining halibut or sablefish with
a salmon permit (Fig. S\ref{fig:ifq}B).

# Discussion

We fit a hierarchical variance-function regression model to 30 years of
commercial fisheries revenue data in Alaska to simultaneously estimate the
effect of changes in species diversity on individual's revenue and revenue
variability. We find that individuals fishing permit sets that allow high levels
of species diversity tend to have lower revenue variability. Adding an
additional fisheries permit usually increases revenue and decreases revenue
variability. However, within the constraints of a permit and accounting for
effort, specializing (reducing species diversity) tends to increase revenue and
specializing or generalizing tends to be associated with greater revenue
variability, with considerable differences across fisheries. Thus, the most
effective mechanism by which species diversity can lower revenue variability for
individual fishers --- switching or adding fishing permits --- is often costly,
limited by regulations, and unavailable to many individuals.

@@

>>>>

# Abstract
Ole would drop the confidence bounds for the abstract.  I added a bit but this seems like a long abstract for PNAS...

A fundamental challenge for individuals who rely on the extraction of natural
resources for their livelihood is income variability. Fishers, loggers, farmers,
and miners, among others, face high levels of income uncertainty driven by a mix
of environmental, biological, and economic variability. A key component
underlying the sustainability of these industries is identifying both how
regulatory actions and individual behaviour affect income variability and by
extension the economic stability and sustainable use of natural resources. For
commercial fisheries, recent work has identified that targeting a diversity of
species can reduce income variability. However, any diversity-variability
relationship is embedded in a regulatory framework that restricts individual
ability to change fishing behaviour. Here we identify three primary axes by
which fishers in Alaska, one of the world's most productive fishery regions, can
potentially diversify fishing activities and investigate how individual actions
along each diversity axis affect both revenue and revenue variability. We
construct a novel statistical model to simultaneously assess the value to
individual fishers of choosing to (1) own permits that allow targeting a diverse
fish species assemblage, (2) diversify by owning multiple permits, or (3) catch
a greater species diversity within the constraints of a given permit. We show
contrasting effects of diversity on revenue and revenue variability along the
three axes. First, we show that fishing permits with higher species diversity
reduces variability (`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction
per unit increase in mean effective species diversity). Next, we show that
fishing an additional permit is associated with higher revenue and lower
variability (e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`%
IQR] decline in standard deviation for fishers with halibut and another permit vs.
only a halibut permit). Finally, we show that increasing species diversity
within the constraints of a permit set has a fishery-dependent effect on revenue
and is usually (`r g2_dens`% probability) associated with increased uncertainty
in revenue the following year. Thus, while there are clear benefits to some axes
of diversification, others are less effective. Importantly, the most effective
mechanism for increasing revenue and decreasing risk --- purchasing additional
permits to participate in additional fisheries --- was also the most costly.  Furthermore, 
the least costly option of increasing catch diversity within an existing permit
was associated with increased revenue variability. Thus, over the past 30 years as
the Alaskan commercial fishing fleet has become increasingly specialized and most individuals 
only participate under a single permit, the practical benefits accrued
from increased species diversification are limited.

Our results suggest that thebenefit of species diversity for revenue variability is mostly an across-permit
effect, not a benefit that can be easily harnessed by an individual permit
holder. Effective methods of diversification --- purchasing additional permits
or switching permits --- are costly and often limited by regulations such as
limited entry programs and individual transferable quotas (ITQs). Take home ...

@@

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
obtainable by commercial fishers [@kasperski2013], possibly because both fish
and fishers are spatially mobile 
[Ole is pretty sure it is because they don't have sufficient political clout, suggest deleting the clause after the comma in previous sentence]. 
However, fishers maintain the ability to
lower risk by diversifying their fishing portfolio. They can, for example, target as many species as permitted by
a license, own multiple licenses, or diversify spatially and fish in multiple
regions. The effectiveness of these diversification strategies for increasing
revenue and reducing revenue variability remains uncertain. For example, theory
and empirical studies suggest there can be substantial benefits to
specialization for both farmers and fishers [@abdulai2000; @arnason2005].
Furthermore, the majority of research examining the effects of diversification
on revenue do so at aggregate levels that combine individuals (e.g. households,
counties, or communities; [@kasperski2013; @sethi2014]). Thus, key
uncertainties remain about the fundamental relationship between diversity,
fishing revenue, and fishing revenue variability at the level of the individual.

@@

Using a unique dataset collected over 30 years (1985--2014) on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals we can track
people that switch fishery permits over time or use multiple vessels. Similarly,
we can disentangle fishers that share the same vessel. We develop and apply
a hierarchical Bayesian model to understand the effect of fished species
diversity [OLE: Call this diversity of catch rather than fished?] on individual 
revenue and revenue variability that builds on results
from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
[@sethi2014] variability.

We develop a variance-function regression model [@davidian1987; @hall1989],
to simultaneously models changes in the mean and variance of individual
revenue as a function of covariates. [OLE: I think we leave this sentence for the methods somewhere- Whereas traditional hierarchical linear models assume homoscedasticity --- that the residual variance does not vary
systematically --- the variance-function regression model allows the residual
variance to be heteroscedastic and vary as a function of predictors.] 
We simultaneously estimate the effect of species diversity (the
inverse of Simpson's diversity index [@simpson1949; @jost2006]) on individual
revenue and revenue variability, and estimate effects both within and across
fishing strategies while controling for fishing effort and latent time effects 
(e.g. changing environmental conditions). To our knowledge, this is the first application of this type of
model in any natural resource field.

@@

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given permit set is more variable (Fig.
\ref{fig:within}). After accounting for fishing effort, specializing tends to be
associated with increased revenue (Fig. \ref{fig:within}; Fig. S\ref{fig:fe},
Fig. S\ref{fig:re}) but also increased variability. On average, there is a ~0.5% increase in the year-to-year
revenue ratio per 1% decrease in species diversity, and across strategies, `r
g2_dens`% of the probability density supports a positive relationship between
specialization and revenue.

However, adjusting (increasing or decreasing) an
individual's fished species diversity within the constraints of a permit set
tends to be associated with greater expected variability or uncertainty in
year-to-year revenue (Fig. \ref{fig:within} strategies above 0; Figs.
S\ref{fig:fe}, \ref{fig:re}, \ref{fig:mirror}). This effect was strongest for
individuals with a halibut permit or halibut and sablefish permits. Individuals
with these strategies experienced ~2% increase in variability for a 1% change in
the year-to-year ratio of species diversity (Fig. \ref{fig:within}). Interestingly, becoming more generalized in species diversity also leads to increased revenue variability (FIG SUPP), suggesting that the act of changing strategies is a contributor to increased variability of revenue.

Levels of fished species diversity above one for these particular permits are primarily
due to bycatch.

[OLE: I would just say - see supplement for discussion of consequences of generalizing and not include this section in the main text]
The converse is also usually true --- generalizing
tends to be associated with lower expected revenue after accounting for fishing
effort (i.e. Fig. S\ref{fig:mirror}A and S\ref{fig:mirror}B are near mirror
images left to right). This benefit of specializing is most apparent after
accounting for fishing effort. Given that fishing for more species tends to be
associated with increased fishing effort (fishing more days), specializing tends
to be ...TODO (Fig. S\ref{fig:noeffort}).

Alaskan halibut and sablefish fisheries were converted to individual fishery
quota (IFQ) management in 1994, which potentially confounds our estimates of the effect of
diversity and a major regulatory change. We refit our models allowing all strategy-level effects
to vary separately before and after the introduction of IFQs for strategies involving halibut
and sablefish permits. The introduction of IFQs reduced variability by ~25% (XX--XX 95% CI)
for individuals with only halibut or sablefish permits (Fig. S\ref{fig:ifq}A). 
Furthermore, for most strategies, the association changing species diversity with increased
revenue variability became stronger after the introduction of IFQs (Fig.
S\ref{fig:ifq}B). This was especially true for individuals fishing only halibut
or sablefish and less so for individuals combining halibut or sablefish with
a salmon permit (Fig. S\ref{fig:ifq}B).

# Discussion
[Ole thinks we should start bigger here to reflect the big main themes in the intro]

For individuals deriving livelihoods from natural resources, theory suggests that adopting strategies that utilize
a diversity of assets [species?] provide substantial benefits over strategies with lower diversity [REFS].
However, the ability to effectively diversify in practice is limited by regulation and market forces.
Here we document the effect of changing species diversity on revenue and revenue variability across 
three scale. Each scale reflects individual level choices that produce varied levels of species diversity.  
We find that individuals fishing permit sets that allow high levels
of species diversity tend to have lower revenue variability. Adding an
additional fisheries permit usually increases revenue and decreases revenue
variability. However, within the constraints of a permit and accounting for
effort, specializing (reducing species diversity) tends to increase revenue and
specializing or generalizing tends to be associated with greater revenue
variability, with considerable differences across fisheries. Thus, the most
effective mechanism by which species diversity can lower revenue variability for
individual fishers --- switching or adding fishing permits --- is often costly,
limited by regulations, and consequently unavailable to many individuals in practice.

Our results shows additional support for the idea that individual fishers can reduce 
their income variability by diversifying their catch portfolios. However, most individuals
don’t appear to change their catch diversity significantly from year to year, and 
most strategies adopted by Alaskan fishermen lie in a high-risk, high-reward quadrant
in which reducing diversification increases both revenue variability and expected revenue (Fig. 4).
Previous work has suggested that changes to fisheries management such as limited entry or
catch share programs may inadvertently act as barrier to diversification and promote 
specializing (Hilborn et al. 2003; Kasperski and Holland 2013). While some of these 
specialized fishing strategies may be associated with increased revenue variability,
the benefit may be that they are also associated with increased revenue.
External market forces may interact in concert with management regulations to promote 
specialization for fishers or other individuals dependent on natural resources. 
In terrestrial agriculture for instance, there are also many instances of specializing 
being profitable; drivers include the ability to specialize on high-profit cash crops
(tobacco, cocoa), the availability of government agricultural subsidies (soybeans, corn in the USA),
and farming on a large industrial scale. With increasing environmental variability and 
extreme events in a changing climate, it is particularly important that individuals who 
rely on natural resources as a primary source of income have strategies to reduce risk.
Though variability in individual revenues may not be actively incorporated into many management
decisions, analyses like ours help to understand the range of diversification strategies 
available and how their benefits may change over time.

Our work compliments previous results that have examined the effects of species diversity at the 
community (Ref) and vessel(Ref) level that show THIS (Sethi, Kasperski, others)

@@

We fit a hierarchical variance-function regression model to 30 years of
commercial fisheries revenue data in Alaska to simultaneously estimate the
effect of changes in species diversity on individual's revenue and revenue
variability.

##3 ERIC
